---
title: "Phylogenetic Analyses"
output:
  html_document:
    df_print: paged
---

```{r setup}
knitr::opts_knit$set(root.dir = "/Users/samanthaworthy/Documents/GitHub/Sandhills-Flora")
```

```{r}
library(ggtree)
library(picante)
library(huxtable)
```

### Read in phylogeny
```{r}
phy=read.tree("./Formatted.Data/FullSeq_MAFFT_RAxML_Tree.newick")
```

### Read in community data matrix
```{r}
trait.cdm=(read.csv("./Formatted.Data/Trait.cdm.csv", 
        check.names = FALSE, header = TRUE, row.names = 1))
```

### Prune the outgroup prior to analyses
```{r}
prune.tree=drop.tip(phy,"HS_05")
```

### Combine phylogeny and community data to check for mismatches
```{r}
Com=match.phylo.comm(prune.tree, trait.cdm)
```

### Extract phylogeny data from combined/consolidated data
### 173 species in each dataframe
```{r}
phy.e=Com$phy
com.e=Com$comm
```

### Convert phylogeny to cophenetic distance
```{r}
phydist=cophenetic(phy.e)
```

#### Separate analyses for each trait
```{r}
life.form.cdm = com.e[c(2,3),]
forest.strat.cdm = com.e[c(4,5,6),]
life.history.cdm = com.e[c(7,8),]
```

### Mean Pairwise Distance for each trait
```{r}
mpd.life.form = ses.mpd(life.form.cdm, phydist, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mpd.forest.strata = ses.mpd(forest.strat.cdm, phydist, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mpd.life.history = ses.mpd(life.history.cdm, phydist, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mpd.life.form
mpd.forest.strata
mpd.life.history
```

### Mean Nearest Taxon Distance for each trait
```{r}
mntd.life.form = ses.mntd(life.form.cdm, phydist, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mntd.forest.strata = ses.mntd(forest.strat.cdm, phydist, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mntd.life.history = ses.mntd(life.history.cdm, phydist, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mntd.life.form
mntd.forest.strata
mntd.life.history
```

### Divide data set into annuals and perennials

```{r}
trait.cdm.t = as.data.frame(t(trait.cdm))
annual.t = subset(trait.cdm.t, trait.cdm.t$Annual == 1)
annual.cdm = t(annual.t)
```

```{r}
perennial.t = subset(trait.cdm.t, trait.cdm.t$Perennual == 1)
perennial.cdm= t(perennial.t)
```

### Combine phylogeny and annual data to check for mismatches
```{r}
annual.com=match.phylo.comm(prune.tree, annual.cdm)
```

### Combine phylogeny and perennial data to check for mismatches
```{r}
perennial.com=match.phylo.comm(prune.tree, perennial.cdm)
```

### Extract data from combined/consolidated data
### 25 species of annual plants in each dataframe
```{r}
phy.annual=annual.com$phy
com.annual=annual.com$comm
```

### Extract data from combined/consolidated data
### 147 species of perennial plants in each dataframe
```{r}
phy.perennial=perennial.com$phy
com.perennial=perennial.com$comm
```

### Convert phylogeny to cophenetic distance
```{r}
phydist.annual=cophenetic(phy.annual)
phydist.perennial=cophenetic(phy.perennial)
```

#### Separate analyses for each trait
```{r}
life.form.annual = com.annual[c(2,3),]
forest.strat.annual = com.annual[c(4,5,6),]
life.form.perennial = com.perennial[c(2,3),]
forest.strat.perennial = com.perennial[c(4,5,6),]
```

### Mean Pairwise Distance for each trait for annuals
```{r}
mpd.life.form.annual = ses.mpd(life.form.annual, phydist.annual, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mpd.forest.strata.annual = ses.mpd(forest.strat.annual, phydist.annual, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mpd.life.form.annual
mpd.forest.strata.annual
```

### Mean Pairwise Distance for each trait for perennials
```{r}
mpd.life.form.perennial = ses.mpd(life.form.perennial, phydist.perennial, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mpd.forest.strata.perennial = ses.mpd(forest.strat.perennial, phydist.perennial, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mpd.life.form.perennial
mpd.forest.strata.perennial
```

### Mean Nearest Taxon Distance for each trait for annuals
```{r}
mntd.life.form.annual = ses.mntd(life.form.annual, phydist.annual, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mntd.forest.strata.annual = ses.mntd(forest.strat.annual, phydist.annual, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mntd.life.form.annual
mntd.forest.strata.annual
```

### Mean Nearest Taxon Distance for each trait for perennials
```{r}
mntd.life.form.perennial = ses.mntd(life.form.perennial, phydist.perennial, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mntd.forest.strata.perennial  = ses.mntd(forest.strat.perennial, phydist.perennial, null.model="taxa.labels", abundance.weighted=FALSE, runs=999, iterations=1000)
mntd.life.form.perennial
mntd.forest.strata.perennial
```




